Community Detection Methods for Generalized Random Dot Product Graphs

### 2021-05-28 proposal/slides.Rmd
added citation for ssc

2. Popularity Adjusted Block Model
    a. Connecting the PABM to the GRDPG
    b. Subspace Clustering for Community Detection
    c. Orthogonal Spectral Clustering

@@

**Example 1**: If every entry $P_{ij} = \theta$, then $A \sim \BG(P)$ is an Erdos-Renyi graph.

@@

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z))
```

@@

1. Note that $A \sim \BG(P)$ is a RDPG since $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$.

@@

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.

@@

**Remark 1** (orthogonality of subspaces):
If $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $x_i^\top x_j = 0$.

@@

**Algorithm**: Sparse Subspace Clustering:

>>>>

2. Popularity Adjusted Block Model
    a. Connecting the PABM to the GRDPG
    b. Orthogonal Spectral Clustering
    c. Sparse Subspace Clustering

@@

**Example 1**: If every entry $P_{ij} = \theta$, then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
For this model, $A_{ij} \iid \Bernoulli(\theta)$.

@@

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

@@

1. Note that $A \sim \BG(P)$ is a RDPG since $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
noting that as $n \to \infty$, the ASE approaches point masses.

@@

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.

@@

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix}$$

$$\Lambda^{(k)} = \begin{bmatrix} 
\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
\end{bmatrix}$$

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix}$$

\vspace*{2\baselineskip}

$$U \in \mathbb{O}(K^2)$$

$$A \sim \PABM(\{\lambda^{(kl)}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$

**Remark 1** (orthogonality of subspaces):
If $y_i^\top$ and $y_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $y_i^\top y_j = 0$.

@@

**Algorithm**: Sparse Subspace Clustering (Elhamifar and Vidal, 2009):
